Hypertension and microbiome

### 2019-04-29 articletwo.Rmd
Salt model

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(microbiome)
library(finriskmetagcommon)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(ggpmisc)
library(Maaslin)
library(parallel)
library(officer)
library(flextable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
```

@@

>>>>

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(finriskmetagcommon)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(ggpmisc)
library(Maaslin)
library(parallel)
library(officer)
library(flextable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
```

@@

```{r lacto data frame}
lacto.abund <- subset_taxa(pseq.genus, Genus == "Lactobacillus") %>%
    abundances(., transform = "clr") %>%
    t %>%
    as.data.frame %>%
    tibble::rownames_to_column("id") %>%
    setNames(tolower(replace.brackets(names(.))))

lacto.full <- meta(pseq.genus) %>%
    tibble::rownames_to_column("id") %>%
    full_join(., lacto.abund, by = "id") %>%
    dplyr::mutate(fake_salt = relevel(factor(dplyr::ntile(BL_AGE, 4)), ref='1'),
                  lacto_groups = relevel(factor(dplyr::ntile(lactobacillus.bacteria, 4)), ref='1'))
```

```{r lacto adjusted means}
lacto.lm.salt <- lm(lactobacillus.bacteria ~ fake_salt + BL_AGE + SEX, data=lacto.full)

lacto.mean.unadjusted <- data.frame(
    fake_salt = as.factor(seq(4)),
    SEX = as.factor(0),
    BL_AGE = mean(lacto.full %>% pull(BL_AGE)))

lacto.mean.adjusted <- predict(lacto.lm.salt, lacto.mean.unadjusted, interval = "confidence") %>%
    as.data.frame %>%
    dplyr::rename(estimate = fit,
           conf.low = lwr,
           conf.high = upr) %>%
    tibble::rownames_to_column("term")
```

```{r lacto odds for hypertension}
lacto.odds <- glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX,
                  family = binomial(link=logit),
                  data=lacto.full) %>%
    broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
    filter(!grepl("Intercept", term)) %>%
    select(term, estimate, conf.low, conf.high) %>%
    add_row(term = "lacto_groups1", estimate = 1.0, conf.low = NA, conf.high = NA) %>%
    dplyr::filter(grepl("lacto", term))
```

```{r lacto lm residuals}
ggplot2::autoplot(lm(lactobacillus.bacteria ~ fake_salt + BL_AGE + SEX, data=lacto.full))
```

```{r lacto logistic residuals}
ggplot2::autoplot(glm(HYPERTENSION ~ lacto_groups + BL_AGE + SEX, family = binomial(link=logit), data=lacto.full))
```

```{r lacto anova}
lacto.anova <- aov(lactobacillus.bacteria ~ fake_salt, data = lacto.full)
summary(lacto.anova)
TukeyHSD(lacto.anova)$fake_salt %>% kable
```

```{r lactoto test normality}
grid.arrange(ggqqplot(lacto.full$lactobacillus.bacteria, title = "Lactobacillus"),
             ggqqplot(lacto.full$BL_AGE, title = "BL_AGE"), ncol=2)
```

```{r forest plot lacto}
myforestplot <- function(data, xlab = "", ylab = "", ylimits = c(0,10), interceptone = FALSE, subtitle = "") {
    ggplot(data = data,
           aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
        geom_point() +
        { if (interceptone) geom_hline(aes(fill=fake_salt), yintercept=1, linetype=2) } +
        geom_errorbar(width=0.1)+ 
        ggplot2::theme_minimal()  +
        theme(plot.title=element_text(size=16),
              axis.title=element_text(size=12),
              strip.background = element_blank()) +
        theme(panel.border = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              axis.ticks.x = element_line(),
              axis.ticks.y = element_line(),
              axis.line = element_line(colour = "black")) +
        scale_x_discrete(labels=c("Q1", "Q2", "Q3", "Q4")) +
        scale_y_continuous(limits = ylimits, breaks = pretty(ylimits, 4)) + 
        xlab(xlab) +
        ylab(ylab) +
        labs(subtitle = subtitle)
}

gsalt <- grid.arrange(
    myforestplot(lacto.mean.adjusted,
                 xlab = "Dietary salt",
                 ylab = "Lactobacillus abundance",
                 subtitle = "A.",
                 ylimits = c(2.5,3)),
    myforestplot(lacto.odds,
                 xlab = "Lactobacillus abundance",
                 ylab = "Odds of hypertension",
                 subtitle = "B.",
                 ylimits = c(0.5, 1.5),
                 interceptone = TRUE),
    ncol=2)
ggsave(file = "cache/salt-forest.png", plot=gsalt, height=4, width=6, dpi = 300)
```

```{r Write figure 3}
fig3head <- "Associations between dietary salt, lactobacillus abundance, and odds for hypertensoin"
fig3foot <- "In panel A we see lactobacillus abundances in different dietary salt groups. The mean value is adjusted using age and sex. In panel B we split paticipants in four groups by lactobacillus abundances and calulate odds ratio for hypertension. Age and sex are included in logistic model. Both panel show  95 % CI for respective variables."
writeimage(doc, 3, "cache/salt-forest.png", fig3head, fig3foot)
```
